Parameterless Population Pyramid

### 2013-12-03 CGP Journal/response.txt

First, we would like to thank both the reviewers and the editorial staff for their
time in examining our work.  The comments we received were very helpful and
have lead to significant improvements.

Before covering individual points, we would like to respond to the two
reviewers who suggested we could improve the support of our conclusions by
including problems outside the binary domain.  We completely agree, and as
with Reviewer 3 we feel like this is an excellent avenue for future work.  However,
the focus of this work was to delve deeply into a small set of problems with
the hope of developing the hypotheses about CGP's inherent behavior.  To that
end we felt attempting to increase the breadth would limit our depth.  As was
pointed out by two reviewers, our work required extensive experimentation
to achieve the conclusions on our small set of problems.  More directly, to
include a single new problem (boolean or otherwise) with which we
could perform comparable analysis would require over four thousand additional runs
to perform proper parameter tuning and data collection.

The reason we chose all of our original problems as binary was to allow for the
detailed analysis of behavior in the latter half of Section VI.  If we had done
four problems, but using diverse domains, it would have been much more challenging
to determine why we observed changes in behavior between problems.  By using
four problems with similar domains and the same operators, we are able to determine
how our techniques scale with problem difficulty.  Similarly, the analysis tools
we developed to perform the comparison could rely on problem uniformity.
Also, since the binary domain is
the most studied and successful in CGP literature, detailed analysis and conclusions drawn about
its behavior in this domain are likely to have the most impact on the field.

A major focus of our work (ordering effect on the location and number of active genes)
is grounded in theory which makes no assumptions on the type of problem being solved.
Furthermore, the operators we investigate deal with domain independent problems.
As the experimental results matched the theory, we have no reason to suspect our
conclusions on these topics to change when the operators or data type is changed.
However, we understand that the following could change depending on data type: The relative effectiveness
of the proposed modifications (IE, skip vs single, reorder vs normal, etc), the
evolved genome sizes, and the charts in Table IV and Table V.  Yet we do not feel
these are the primary concern of the work, nor do we have any reason to believe
the conclusions should only hold true for binary problems.

With that said, we would now like to provide point by point responses to the reviewer
comments.

Reviewer 1:
"You may find it useful to at least mention other representations of CGP in
which the dependency of node ordering has been reduced, or even, removed"
Page 3, right column, last paragraph before Section Reorder: This paragraph was
added with appropriate citations to a few versions of CGP which do not follow Equation 1.

"When comparing your results with previous publications [8], why didn't you used the same function set?"
In [8], the authors use AND, OR, NAND, NOR for all of the other boolean problems tested (Even Parity, Digital Adder, and Digital Comparitor),
but use AND, AND with one input inverted, XOR and OR on digital multiplier.
The two papers [8] cites that performed research on this problem used AND, XOR.  They claim this makes the problem easier as
human designed solutions use only AND, XOR.  Therefore eliminating XOR as a choice likely makes this problem even harder.
That said, our paper chose to keep the operators consistent across problems.  Converting our
version to use their operators (either [8]s or the work cited by [8]) would require an infeasible
amount of time to again perform tuning.  However, Page 8, left column we have added more
information about this comparison.  Notably, we used our parameters with their operators
and maintained Reorder as being statistically better.

"It is unfortunate that all analysis was limited to boolean problems"
Covered in general response.

"[I]t may be helpful to make it clear that Skip, Accumulate etc. are your invention"
For terseness we felt that as our abstract states our focus was on "creating methods"
combined with fact that these methods both exist in previous published work, no modification
was necessary.

"[I]n many CGP implementations it is not common to exclude the gene's current value during mutation."
Page 1, right column, last paragraph: We have correct this to point out both that it isn't
so much common as unclearly defined and that our choice is in doing so is to allow
for precise behavioral analysis.  While the distinction may result in minor differences in behavior,
we see no reason for it to effect any conclusions drawn about CGP's aggregate behavior.

"Once all of a *node's* dependencies..."
Fixed.

"References or at least a very brief description of the purpose of for
Kruskal–Wallis (as for the Mann-Whitney U test) would aid those not familiar."
Page 7, Footnote 3 and 4 have been added.

Reviewer 2:
"Please provide the reasons why the interaction is important? Also, please explain (briefly) how your goals can be achieved."
Page 1, left column, last sentence as been added.

"[W]hy only "one" active gene is mutated?"
Page 3, left column, last paragraph before Section Genome Reordering: Added explanation that more than
one could have been used, but one is sufficient and removes a parameter.

"", and defined Normal as CGP's historical ordering method", do you mean the ordering method of the original CGP?"
Yes.  Page 3, right column, paragraph starting with "To rectify these", last sentence clarifies our meaning.

""While this increase ... drawback.", this sentence is not entirely clear to me. Please elaborate more and provide reasons for this statement."
Page 4, right column, last paragraph before Section DAG.  Last few sentences expand on the idea.

""mutating node" or "mutated node"?"
Mutating is in fact intended here.  This section is discussing the process of performing
a mutation on a node, so at this point in the description it is "mutating."

"It would be nice to give some rationales why potential configurations run 5 times and later 4 times? Why are higher values used?"
Page 6, right column, last full paragraph, starting from "We chose to run each initially 5 times"

""we would suggest Skip be used over accumulate for any real world application." is quite a strong claim. Only binary problems are examined in the experiments, is it enough to support your claim?"
Page 7, right column, last full paragraph: Added some language to clarify our reasoning.  From what we can tell
the methods are different ways of doing the exact same thing.  Since Accumulate is more complicated, we suggest Skip.

"If possible, please provide the average, standard deviation of "Evaluations to success" and the average running time of each configuration.
We further examined the statistical measures
of evaluations to success which we report in Table II.  As it is not safe to assume the number
of evaluations to success is normally distributed (in fact, they are most likely not normal),
we chose to provide non-parametric (median) instead of parametric (mean) measures,
as was done in [8].  Upon our review, we decided Median Absolute Deviation was not
capturing useful information, and we now show bootstrap confidence intervals, which is a much
more statistically sound and interesting value.

As to the including the wall clock time for our experiments, we attempted to cover our
reasoning for not doing this starting on Page 8, right column, continuing to the end of the section.
In order to provide a fair comparison of runtime, we would need to redo tuning
with a focus on minimizing wall clock time instead of number of evaluations, which
would require another four thousand runs to time just one problem.  An additional
problem in reporting this value is hardware diversity.  To complete our experiments
we utilized a computing cluster with diverse hardware quality.  To avoid this problem,
we would need to recreate all experiments on homogeneous hardware, which is intractable
for experiments of this size.

Reviewer 3:
"Although I am not suggesting any more work be undertaken. It would have been nice
to see more work done on harder problems. ... The paper has focused on Boolean problems and it would be interesting to
carry out similar analysis in other domains. Of course, this remains for the future."
Covered in general response.

"Mutation is generally carried out in two ways [rate vs probability]...The two methods require markedly different amounts of computation"
It is possible to achieve O(1) amortized cost when performing mutations using the probability method.
This can be done leveraging the binomial distribution, which in many languages is implemented to give
the random number of trials k of N total which are true given probability P.  k genes can then
be selected from the genome at random just as when using a mutation rate.  This operator is
advantageous in that allows for a flexible number of mutating genes, up to and including
mutating all genes in a single step.  We alluded to this Page 9, left column, first full paragraph.
Note, if we were to extend Single to use a configurable
number of genes, this would be my choice of method.  As such we would disagree that the probability method
by necessity requires more computation than rate.

That said, for clarity we have corrected all references to mutation "rate" and replaced them with "probability."

"Another quite minor point is that as stated in Miller's book (page 28) when a function gene is selected for mutation, no
check is made as to whether the mutated function is the same as that before mutation"
Page 1, right column, last paragraph: We have clarified this.  See response to similar comment by Reviewer 1.

"than" should be "then"
Fixed.

"It is worth noting in the paper that this step does incur a computational cost."
Page 2, left column, last full paragraph.

"I think the paper would benefit from some
remarks about the extra computational cost of carrying out the Accumulate strategy"
Page 2, right column, last paragraph.

"Since one knows what the parent's active genes are, one could change an active gene at random."
Page 3, left column, last paragraph before Section Genome Ordering partially covers this.  In general
we felt the existing arguments about the importance of neutral drift covered why inactive genes are
also mutated.

"...this representation likely..." -> "...this representation is likely..."
Fixed.

"convenience" -> "convenient"
Page 4, left column, first paragraph: We decided the word convenient was not necessary and removed
it entirely.

"effecting" -> "affecting"
Fixed.

"It is probably a Latex issue, but the caption TABLE 1 merges with the bottom line of the table."
The IEEE standard is to have captions above a table.  Our original submission had captions below
the tables, which was why there was a problem.  Caption locations for all tables have been corrected.

"making comparisons with modular CGP on 3-bit parity is of little value actually"
That comparison has been removed as we felt it was not useful.

"the are" -> "they are"
Fixed.

"If this is true, it may help to explain CGP uses such large genomes..."
Fixed.

"most effect" -> "Largest effect"?
Fixed.

### 2013-12-05 CGP Journal/response.tex

\documentclass{letter}
\usepackage{hyperref}
\usepackage{verbatim}
%\signature{Brian W. Goldman and William F. Punch}
%\address{21 Bridge Street \\ Smallville \\ Dunwich DU3 4WE}
\begin{document}

\begin{letter}{}
\opening{Greetings Editors and Reviewers:}

@@

First, we would like to thank both the reviewers and the editorial staff for their
time in examining our work.  The comments we received were very helpful and
have lead to significant improvements.

Before covering individual points, we would like to respond to the two
reviewers who suggested we could improve the support of our conclusions by
including problems outside the binary domain.  We completely agree, and as
noted by Reviewer 3 we feel like this is an excellent avenue for future work.  However,
the focus of this work was to delve deeply into a small set of problems to
develop hypotheses about CGP's inherent behavior.  To that
end we felt that increasing the paper's breadth would limit our depth.  As was
pointed out by two reviewers, our work required extensive experimentation
to achieve our conclusions on a small set of problems.  More directly, to
include a single new problem (boolean or otherwise) with
comparable analysis would require over four thousand additional runs.
%including proper parameter tuning and data collection.

The reason we chose binary problems was to allow for the
detailed analysis of behavior in the latter half of Section VI.  If we had done
four problems, but using diverse domains, it would have been much more challenging
to determine why we observed changes in behavior between problems.  By using
four problems with similar domains and the same operators, we were able to determine
how our techniques scale with problem difficulty.  Similarly, the analysis tools
we developed to perform the comparison allowed us to rely on problem uniformity.
Finally, since the binary domain is
the most studied and successful in CGP literature, detailed analysis and conclusions drawn
are likely to have the most impact on the field.

We would like to reiterate that our thesis is understanding
CGP's problem solving behaviors and not necessarily to suggest any of our modifications
as universal improvements.  We agree that the inclusion of wall time comparisons on a diverse
set of hard real world problems can be very informative when defending new methods
for optimization, but we strive more for comprehension than improvement.

\begin{comment}
A major focus of our work (ordering effect on the location and number of active genes)
is grounded in theory which makes no assumptions on the type of problem being solved.
Furthermore, the operators we investigate deal with domain independent problems.
As the experimental results matched the theory, we have no reason to suspect our
conclusions on these topics to change when the operators or data type is changed.
However, we understand that the following could change depending on data type: The relative effectiveness
of the proposed modifications (IE, skip vs single, reorder vs normal, etc), the
evolved genome sizes, and the charts in Table IV and Table V.  Yet we do not feel
these are the primary concern of the work, nor do we have any reason to believe
the conclusions should only hold true for binary problems.
\end{comment}

@@

"When comparing your results with previous publications [8], why didn't you used the same function set?"
In [8], the authors use AND, OR, NAND, NOR for all of the other boolean problems tested (Even Parity, Digital Adder, and Digital Comparitor),
but use AND, AND with one input inverted, XOR and OR on digital multiplier.
The two papers [8] cites that performed research on this problem used AND, XOR.  They claim this makes the problem easier as
human designed solutions use only AND, XOR.  Therefore eliminating XOR as a choice likely makes this problem even harder.
That said, our paper chose to keep the operators consistent across problems.
However, Page 8, left column we have added more
information about this comparison.  Notably, we used our parameters with their operators
and maintained Reorder as being statistically better.

"It is unfortunate that all analysis was limited to boolean problems"
Covered in general response.

"[I]t may be helpful to make it clear that Skip, Accumulate etc. are your invention"
For terseness we felt that as our abstract states our focus was on "creating methods"
combined with fact that these methods both exist in previous published work, no modification
was necessary.

"[I]n many CGP implementations it is not common to exclude the gene's current value during mutation."
Page 1, right column, last paragraph: We have corrected this to point out both that it isn't
so much common as unclearly defined and that our choice is in doing so is to allow
for precise behavioral analysis.  While the distinction may result in minor differences in behavior,
we see no reason for it to effect any conclusions drawn about CGP's aggregate behavior.

@@

"It would be nice to give some rationales why potential configurations run 5 times and later 4 times? Why are higher values used?"
Page 6, right column, last full paragraph, starting from "We chose to run each initially 5 times"

""we would suggest Skip be used over accumulate for any real world application." is quite a strong claim. Only binary problems are examined in the experiments, is it enough to support your claim?"
Page 7, right column, last full paragraph: Added some language to clarify our reasoning.  From what we can tell
the methods are different ways of doing the exact same thing.  Since Accumulate is more complicated, we suggest Skip.

"If possible, please provide the average, standard deviation of "Evaluations to success" and the average running time of each configuration.
We further examined the statistical measures
of evaluations to success which we report in Table II.  As it is not safe to assume the number
of evaluations to success is normally distributed (in fact, they are most likely not normal),
we chose to provide non-parametric (median) instead of parametric (mean) measures,
as was done in [8].  Upon our review, we decided Median Absolute Deviation was not
capturing useful information, and we now show bootstrap confidence intervals, which is a much
more statistically sound and interesting value.

As to the including the wall clock time for our experiments, we attempted to cover our
reasoning for not doing this starting on Page 8, right column, continuing to the end of the section.
In order to provide a fair comparison of runtime, we would need to redo tuning
with a focus on minimizing wall clock time instead of number of evaluations.
An additional
problem in reporting this value is hardware diversity.  To complete our experiments
we utilized a computing cluster with diverse hardware quality.

@@

"convenience" -> "convenient"
Page 4, left column, first paragraph: We decided the word convenient was not necessary and removed
it entirely.

"effecting" -> "affecting"
Fixed.

"It is probably a Latex issue, but the caption TABLE 1 merges with the bottom line of the table."
The IEEE standard is to have captions above a table.  Our original submission had captions below
the tables, which was why there was a problem.  Caption locations for all tables have been corrected.

"making comparisons with modular CGP on 3-bit parity is of little value actually"
That comparison has been removed as we agree it was not useful.

### 2013-12-13 Parameterless Population Pyramid/main.tex
Added table and figure, more filled in sections.

@@

\subsection{All Together}
\label{sec-alltogether}
TODO Pieces working together.  HC removes pairwise noise, which allows crossover to detect
clusters.  Levels allow realization of linkage at high levels, reduction in noise.

@@

LTGA works by iteratively improving a population of solutions in a generational manner.
Each generation the half of the solutions (chosen using standard tournament selection)
are used to construct a linkage tree just as described in Section~\ref{sec-crossover}.
Each solution in the population is then crossed with the entire population using that
linkage tree, with the results put into the next generation.  This process continues
until the population fails to change between generations, signaling convergence.

Unlike P3, LTGA does not make use of a hill climber, and there is evidence that
doing so results in decreased efficiency~\cite{bosman:2011:lsbbo}.  LTGA applies
clusters in least linked first order whereas P3 uses smallest first.  Also, LTGA
does not search for donors until one containing different genes is found.  Instead
if the donated genes are identical, it simply skips the evaluation.  LTGA
requires a population size parameter, maintains a fixed population size across
generations, and does not prevent duplicate individuals in a population.  LTGA
does not add new genetic information after initialization, meaning that if
required diversity is not contained in the original population or lost in future
generations, LTGA will fail to find the global optimum.

@@

We also made minor modifications to selection and to prevent wasting evaluations.
If the best offspring produced in a generation is a mutant offspring, select that
over any of those produced using crossover.  If there is a fitness tie between the
best offspring in a generation, select the one with the maximum hamming distance
to the parent.  This encourages drift over plateaus.  While the original paper
discusses a ``mod'' version of the $\lambda$ control strategy for how to handle
when offspring of equal fitness are produced, we found this conflicted with our
method of restarting.  As such we use the original control strategy, where $\lambda$
is increased if the best offspring in a generation is not strictly better than its parent.

\section{Test Problems}
TODO Mention that since all of our algorithms are variable order insensitive
and have no bias toward creating 1s over 0s, all will have identical results if
problem variables are randomly reordered or XOR'd.

TODO Chose only problems with known global optima to allow comparison on time to
global optima.

@@

While both of these problems have difficult to solve subproblems, there non-overlapping
nature and single optimal setting may not capture aspects of real world problem solving.
Therefore the Hierarchical If and only If (HIFF) problem is often used to test an
algorithms ability to handle these characteristics~\cite{thierens:2013:ltgahiff}.
In this problem the genome is broken up hierarchically into a complete binary tree,
with each gene representing a single bit, and each internal node representing the union
between all genes in that nodes subtree.  A node in the tree contributes to the fitness
only if all of the genes in its subset have the same value (IE are all 0 or are all 1).
The amount contributed is equal to the number of bits in the subset.  As a result
this problem involves creating larger and larger blocks of single value runs which
only score if the entire block is optimized.

Our final single instance problem is borrowed from the real valued optimization domain.
The Discretized Rastrigin used the standard Rastrigin evaluation function (Equation~\ref{eq-rast}).
\begin{equation}
  An + \sum_{i=1}^{n}\left [ x_i^2-A\cos (2\pi x_i) \right ] \forall x\in [-5.12,5.12]
  \label{eq-rast}
\end{equation}
This function creates a highly multimodal landscape which is completely separable
and has a single global optima.  In order to apply our algorithms to this complex problem,
we discretized the representation using 10 bit gray coded numbers evenly distributed
on the range $[-5.12,5.12)$.

@@

Ising Spin Glasses are another large class of $NP-Hard$ problems.  These problems
involve minimizing the interaction terms in a graph based on edge weights by changing
the sign of graph nodes.  Similar to with NK landscapes, there is a subset of Ising problems
known as $2D\pm J$ Ising Spin Glasses.  This subset restricts graph interactions to a
toroidal two dimensional grid with edge weights of $\{-1, +1\}$.  This subset can
be solved in polynomial time~\cite{saul:1994:spinglass}, and there are existing
tools for solving arbitrary instances.\footnote{\url{http://www.informatik.uni-koeln.de/spinglass/}}

Our final problem class is a subset of the Maximum Satisfiability (MAXSAT) problem.
This problem consists of a series of disjunctive clauses, such that each clause contains
three problem variables, some of which are negated. While
MAXSAT requires exponential time to solve, we have devised a method of generating
instances that ensures each has at least one solution that satisfies all clauses.
First, a random target solution is generated.  For each clause, set the sign of each
variable in the clause such that that clause will be satisfied by the target solution.
To ensure there is no bias to variable signs in the generated problem there is a $\frac{1}{6}$ probability
all clause signs match the target, a $\frac{1}{6}$ probability two of the three signs match,
and a $\frac{4}{6}$ probability only one sign matches.  By setting the clause signs
in this way we ensure the target solution is always able to satisfy all clauses,
meaning all problems have a known best fitness.  Note that its possible there to be
other solutions that satisfy all clauses.

\section{Experimental Results}
TODO Cover all 7 problems as a group, focusing on the two ``types'' of problems.
Make it explicit that you outperform by a constant factor on the static problems
and by an order of complexity on generated problems.

@@

TODO Include graphic for 7 problem results.

TODO Include table comparing results on largest version of each problem.

>>>>

\newcommand{\includegraphicswide}[1]
{\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{#1}}

@@

\subsection{All Together}
\label{sec-alltogether}
The pieces of P3 were specifically chosen to interact in desirable ways.  First,
the hill climber provides an efficient tool for discovering pairwise linkage, as
search only ends when no single bit change can be made which improves fitness.
Therefore crossover, which relies on pairwise linkage, has high quality information
to leverage when determining clusters.

Similarly, by starting from a random position for each new solution, P3 progressively reduces
the chance of spurious linkage formation due to initialization.  Furthermore, these random
restarts provide exploration to balance out the hill climber and crossover's extreme
exploitation.

Finally, the use of levels allows P3 to perform solution filtering without throwing
away information.  Once found, a local optima is stored indefinitely and its linkage
and genes will have impact on future crossovers.  However, by splitting solutions
into levels, the linkage information and the donated genes come from similar solution
qualities.  This prevents noise in gene linkage which only exists in low quality
solutions from effecting cluster formation for high quality solutions.  It also
reduces the chance for high quality solutions to simply overwrite new, unoptimized
solutions, preserving search diversity.

@@

LTGA works by iteratively improving a population of solutions in a generational manner.
Each generation half of the solutions (chosen using binary tournament selection)
are used to construct a linkage tree just as described in Section~\ref{sec-crossover}.
Each solution in the population is then crossed with the entire population using that
linkage tree, with the results put into the next generation.  This process continues
until the population fails to change between generations, signaling convergence.

Unlike P3, this variant of LTGA does not make use of a hill climber, and there is evidence that
doing so results in decreased efficiency~\cite{bosman:2011:lsbbo}.  LTGA applies
clusters in least linked first order whereas P3 uses smallest first.  Also, LTGA
does not search for donors until one containing different genes is found.  Instead
if the donated genes are identical, it simply skips the evaluation.  LTGA
requires a population size parameter, maintains a fixed population size across
generations, and does not prevent duplicate individuals in a population.  LTGA
does not add new genetic information after initialization, meaning that if
required diversity is not contained in the original population or lost in future
generations, LTGA will fail to find the global optimum.

@@

We also made minor modifications to selection and to prevent wasting evaluations.
If the best offspring produced in a generation is a mutant offspring, select that
over any of those produced using crossover.  If there is a fitness tie between the
best offspring in a generation, select the one with the maximum hamming distance
to the parent.  This encourages drift over plateaus.  While the original paper
discusses a ``mod'' version of the $\lambda$ control strategy for how to handle
when offspring of equal fitness are produced, we found this conflicted with our
method of restarting.  As such we use the original control strategy, where $\lambda$
is increased if the best offspring in a generation is not strictly better than its parent.
If an offspring produced by crossover has identical genes to either of its parents,
it is not evaluated.

\section{Test Problems}
All four of our optimization techniques share some interesting properties for
problem equivalence.  First, all are fitness scale invariant, in that changing
the fitness values without changing the relation between solution fitnesses does
not effect any of their search methods.  This is true because none of the methods
uses the fitness values directly, only when comparing the relative fitness between
two solutions using boolean operators.  This does not hold true for some optimization
methods, for instance fitness proportional selection.

None of the methods have a bias toward creating specific gene values outside
of the bias created by the problem itself.  This means all of the optimizers
will have unchanged behavior if a problem's inputs are reflected or XOR'd with a fixed string.  As
an example, the problem of maximize the number of ones in a solution will be treated identically
to the problem of maximizing the number of zeros in a solution.  Both are also identical
to the problem of minimizing the hamming distance to any arbitrary solution.

All of the methods are gene order independent.  This means that the location
of genes in the genome can be randomly shuffled without effecting search behavior.
This is not true in many evolutionary systems due to crossover.  For example,
when using one point crossover the distance between genes in the genome effects the
probability they will be kept together through a crossover.

Combined, these features mean that tests on apparently simple problems can be
indicative of search behavior on a large class of actually complex problems.

In choosing test problems, we decided to use only benchmarks with knowable global
optima.  This allows us to measure how many evaulations each method requires to converge to the
global optima, as opposed to just the highest fitness reached.  This is advantageous
as three of the four algorithms have no hard termination condition.  The remaining
algorithm, LTGA, uses a population parameter which can be increased to slow
convergence with the potential increase in fitness reachable before convergence.
Therefore it is difficult to say how and when to declare any of the techniques
converged unless the global optimum is found.
While in practice it is often more important to find a good solution quickly
than to find the optimum after an exceedingly long time, using this as a measure
for algorithm quality can lead to algorithms which prematurely converge.

@@

While both of these problems have difficult to solve subproblems, there non-overlapping
nature and single optimal setting may not capture aspects of real world problem solving.
Therefore the Hierarchical If and only If (HIFF) problem is often used to test an
algorithms ability to handle these characteristics~\cite{thierens:2013:ltgahiff}.
In this problem the genome is broken up hierarchically into a complete binary tree,
with each gene representing a leaf, and each internal node representing the union
between all genes in that node's subtree.  A node in the tree contributes to the fitness
only if all of the genes in its subset have the same value (IE are all 0 or are all 1).
The amount contributed is equal to the number of bits in the subset.  As a result
this problem involves creating larger and larger blocks of single value runs which
only score if the entire block is optimized.

Our final single instance problem is borrowed from the real valued optimization domain.
The Discretized Rastrigin problem uses the standard Rastrigin evaluation function (Equation~\ref{eq-rast}).
\begin{equation}
  An + \sum_{i=1}^{n}\left [ x_i^2-A\cos (2\pi x_i) \right ] \forall x\in [-5.12,5.12]
  \label{eq-rast}
\end{equation}
This function creates a highly multimodal landscape which is completely separable
and has a single global optima.  In order to apply our algorithms to this complex problem,
we discretized the representation using 10 bit gray coded numbers evenly distributed
on the range $[-5.12,5.12)$.

@@

Ising Spin Glasses are another large class of $NP-Hard$ problems.  These problems
involve minimizing the interaction terms in a graph based on edge weights by changing
the sign of graph nodes.  Similar to with NK landscapes, there is a polynomially solvable subset of Ising problems
known as $2D\pm J$ Ising Spin Glasses.  This subset restricts graph interactions to a
toroidal two dimensional grid with edge weights of $\{-1, +1\}$.  This subset can
be solved in polynomial time~\cite{saul:1994:spinglass}, and there are existing
tools for solving arbitrary instances.\footnote{\url{http://www.informatik.uni-koeln.de/spinglass/}}

Our final problem class is a subset of the Maximum Satisfiability (MAXSAT) problem.
This problem consists of a series of disjunctive clauses, such that each clause contains
three problem variables, some of which are negated. While
MAXSAT requires exponential time to solve, we have devised a method of generating
instances that ensures each has at least one solution that satisfies all clauses.
First, a random target solution is generated.  For each clause, set the sign of each
variable in the clause such that that clause will be satisfied by the target solution.
To ensure there is no bias to variable signs in the generated problem there is a $\frac{1}{6}$ probability
all clause signs match the target, a $\frac{1}{6}$ probability two of the three signs match,
and a $\frac{4}{6}$ probability only one sign matches.  By setting the clause signs
in this way we ensure the target solution is always able to satisfy all clauses,
meaning all problems have a known best fitness.  Note that its possible there to be
other solutions that also satisfy all clauses.

\section{Experimental Results}
\begin{figure*}
  \centering
  \includegraphicswide{WithRast}
  \caption{Test problems}
  \label{fig-results}
\end{figure*}

\begin{table}
	\centering
	\begin{tabular}{|c|r|c|}
	  \hline
	                & \textbf{Median} & \textbf{Confidence Interval} \\ \hline
    \multicolumn{3}{|c|}{\textbf{Deceptive Trap}, N=805, Speedup=2.936} \\ \hline
\textbf{LTGA} & 1,807,127 & 1,805,039 .. 1,808,718 \\ \hline
\textbf{P3} & 615,503 & 577,000 .. 653,504 \\ \hline\hline
    \multicolumn{3}{|c|}{\textbf{Deceptive Step Trap}, N=805, Speedup=1.107} \\ \hline
\textbf{LTGA} & 14,106,751 & 14,102,987 .. 14,109,910 \\ \hline
\textbf{P3} & 12,742,411 & 11,951,148 .. 13,777,090 \\ \hline\hline
    \multicolumn{3}{|c|}{\textbf{HIFF}, N=2048, Speedup=3.1278} \\ \hline
\textbf{LTGA} & 1,847,077 & 1,845,662 .. 1,848,951 \\ \hline
\textbf{P3} & 590,519 & 570,739 .. 608,160 \\ \hline\hline
    \multicolumn{3}{|c|}{\textbf{Discretized Rastrigin}, N=800, Speedup=1.1854} \\ \hline
\textbf{LTGA} & 244,086 & 242,815 .. 245,189 \\ \hline
\textbf{P3} & 205,908 & 201,012 .. 210,913 \\ \hline\hline
    \multicolumn{3}{|c|}{\textbf{Nearest Neighbor NK}, N=600, Speedup=4.408} \\ \hline
\textbf{LTGA} & 37,620,027 & 37,244,617 .. 37,990,772 \\ \hline
\textbf{P3} & 8,533,322 & 7,315,275 .. 9,481,271 \\ \hline\hline
    \multicolumn{3}{|c|}{\textbf{Ising Spin Glass}, N=784, Speedup=20.403} \\ \hline
\textbf{LTGA} & 12,677,619 & 12,601,917 .. 12,761,798 \\ \hline
\textbf{P3} & 621,347 & 568,100 .. 667,236 \\ \hline\hline
    \multicolumn{3}{|c|}{\textbf{MAX-SAT}, N=60, Speedup=942.50} \\ \hline
\textbf{LTGA} & 21,130,960 & 17,576,126 .. 28,314,680 \\ \hline
\textbf{P3} & 22,420 & 14,363 .. 29,161 \\ \hline

  \end{tabular}
	\caption{Results}
	\label{table-results}
\end{table}

TODO Cover all 7 problems as a group, focusing on the two ``types'' of problems.
Make it explicit that you outperform by a constant factor on the static problems
and by an order of complexity on generated problems.

@@

TODO Compare LTGA results with previously published results as best as possible.
